Verifying Spatial Properties of Array Computations

### 2017-04-18 oopsla17/evaluation.tex
rephrased the UM code mining section

We measured the frequency at which individual specifications involved
multiple occurrences of \term{*}
and \term{+}:
\begin{center}
\vspace{0.25em}
\setlength{\tabcolsep}{0.57em}
{\small{
\hspace{-1em}\begin{tabular}{c|cccccccc}
& 0 & 1 & 2 & 3 & 4 & 5 & 6  \\ \hline
\term{*} & \num{\overallmulOpszero} & \num{\overallmulOpsone} & \num{\overallmulOpstwo} & \num{\overallmulOpsthree} & \num{\overallmulOpsfour} &  & \num{\overallmulOpssix} \\
\term{+} & \num{\overallplusOpszero} & \num{\overallplusOpsone} & \num{\overallplusOpstwo}
\end{tabular}}}
\vspace{0.2em}
\end{center}

@@

One problem we face while examining the millions of line of source
code is that we do not have the original programmer's intention at
hand. But sometimes we have the next best thing, the log entry in the
revision control system. Scanning through the UM we identified an
example of an off-by-1 error that could have been avoided through the
use of stencil specification. The buggy code was identified as
\pointed\ in all dimensions:

@@

One of the initial goals of CamFort was to provide tools that enable
optimisation and refactoring without changing behaviour, and the
array specification feature was designed with that in mind. But
while we were perusing real-world source bases, we realised that there
were also many cases where CamFort could help with refactoring
precisely because they required changes in behaviour. For example, a
relatively common change observed in the logs of revision control for
our corpus is the refactoring of array dimensions. Either
re-ordering, adding or deleting dimensions. Any of these, if not
perfectly propagated throughout all uses, could result in unexpected
outcomes. However, CamFort will pick up the difference during its
stencil specification checking. Ideally, an experienced programmer may
take advantage of the region variable feature of CamFort in order to
minimise the number of specifications that need changing and gain the
full time-saving advantage. But even if not, a quick search for the
specifications can have them updated quickly and then any code that is
not updated will trigger an error when checked.

>>>>

We measured the frequencies of occurrences of the operators \term{*} and \term{+} within stencil specifications inferred from our corpus:
\begin{center}
\vspace{0.25em}
\setlength{\tabcolsep}{0.57em}
{\small{
\hspace{-1em}\begin{tabular}{c|cccccccc}
Occurrences: & 0 & 1 & 2 & 3 & 4 & 5 & 6  \\ \hline
\term{*} & \num{\overallmulOpszero} & \num{\overallmulOpsone} & \num{\overallmulOpstwo} & \num{\overallmulOpsthree} & \num{\overallmulOpsfour} &  & \num{\overallmulOpssix} \\
\term{+} & \num{\overallplusOpszero} & \num{\overallplusOpsone} & \num{\overallplusOpstwo}
\end{tabular}}}
\vspace{0.2em}
\end{center}

@@

Detecting whether an error could have been prevented by a stencil
specification is difficult because it requires knowledge of the
programmer's intent. However, this is occasionally captured in commit
messages accompanying a code change. For example in the Unified Model
we identified an example of an off-by-1 error that could have been
avoided through the use of stencil specification. The buggy code was
identified as \pointed\ in all dimensions:

@@

One of the initial goals of CamFort was to provide tools that enable
optimisation and refactoring without changing behaviour, and the
array specification feature was designed with that in mind. But
while we were perusing real-world source bases, we realised that there
were also many cases where CamFort could help with refactorings
precisely because they required changes in behaviour. For example, a
relatively common change observed in the logs of revision control for
our corpus is the refactoring of array dimensions. Either
re-ordering, adding or deleting dimensions. Any of these, if not
perfectly propagated throughout all uses, could result in unexpected
outcomes. However, CamFort will pick up the difference during its
stencil specification checking. Ideally, an experienced programmer might
take advantage of the region variable feature of CamFort in order to
minimise the number of specifications that need changing and gain the
full time-saving advantage. But even if not, a quick search for the
specifications can have them updated quickly and then any code that is
not updated will trigger an error when checked.
